%-------------------------------------------------------------------------------
\chapter{Introduction}
\label{ch:intro}
%-------------------------------------------------------------------------------

In recent years, \ac{WASM}~\cite{haas2017bringing} has gained prominence~\cite{musch2019new} as a versatile compilation target, serving not only to web-based applications but also to a broader spectrum of use cases~\cite{wasm_use_cases}.
At its core, \ac{WASM} is engineered as an efficient compilation target for high-level, compiled languages such as C and C++.
A fundamental aspect of its design is its linear memory model, which allows these languages to efficiently compile to \ac{WASM} and \ac{WASM} to efficiently compile to a wide variety of architectures.

While WebAssembly provides a sandbox for untrusted code, which protects the host and other guests from malicious or buggy code, it does not inherently prevent memory safety issues within an application's memory space.
This limitation becomes particularly evident when compiling languages like C or C++, where there are no language-level guarantees to prevent these types of bugs.

Recent advancements in hardware, such as ARM's \ac{PAC} and \acf{MTE}, offer promising, high-performance solutions by providing building blocks with low overhead.
These hardware extensions are designed to effectively address memory safety concerns.

We introduce an extension to \ac{WASM} to address both spatial and temporal memory safety bugs with a prototype implementation utilizing \ac{MTE} to implement the safety guarantees efficiently.
Our approach requires no modification to the source code, as we provide a complete compiler toolchain and standard library that can be used to harden unmodified C/C++ programs.
The core contributions of this thesis are outlined as follows:

\begin{description}
    \item[WebAssembly Extension:] A minimal and generic extension to the WebAssembly instruction set, allowing for protected memory regions without code changes.
    \item[Compiler Toolchain:] A compiler toolchain that transparently transforms unmodified programs to provide spatial and temporal memory safety for stack and heap allocations.
    \item[Runtime:] A \ac{WASM} runtime with support for our WebAssembly extension and modified \ac{WASM} compiler that utilizes ARM's \ac{MTE}.
    \item[Bounds Checks with \ac{MTE}:] An implementation to eliminate expensive software-based bounds checks for 64\,bit \ac{WASM} programs.
    \item[Evaluation on real hardware:] Evaluation of our implementation on ARM hardware, including performance and memory overheads as well as security guarantees, and \ac{MTE} as implemented in actual production hardware.
\end{description}
