\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}

\vspace*{20mm}

\begin{center}
{\usekomafont{sectioning}\usekomafont{section} Acknowledgments}
\end{center}

\vspace{10mm}

First and foremost, I would like to thank my supervisor, Pramod Bhatotia.
My academic life would have been quite different without his guidance and support.
Since writing my Bachelor's thesis in 2021, his chair has been a welcoming place, allowing me to grow personally and as a researcher while opening many doors for the future.

\noindent
I am also grateful to my advisor, Dimitrios Stavrakakis, for answering my questions about memory safety, giving me feedback on my work, and helping me debug weird measurements.


To my friends, who have been with me throughout my master's thesis and beyond in both academic and personal journeys:
Thank you for making this time enjoyable.
While I cannot name everyone here, I want to extend special thanks to Andreas and Christoph for their feedback on this thesis.

During my last semester break, I had the opportunity to do an internship at Huawei's Helsinki System Security Lab.
I gained extensive knowledge in compilers, WebAssembly, and security, greatly enhanced by the collaboration with my colleagues Antti, Janne, Carlos, Rémi, Valentin, and my manager, Jan-Erik, despite our short time sharing the office.

Lastly, I'd like to thank my family.
Their support enabled me to study in Munich, and I am forever thankful for the unconditional support they have provided me.

\cleardoublepage{}
